Adaptive Sequential Model Selection

### 2015-05-05 paper.tex
Added more on stopping rules

\newcommand{\cM}{\mathcal{M}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\FDR}{\textnormal{FDR}}
\newcommand{\FCR}{\textnormal{FCR}}
\newcommand{\crt}{\phi}
\newcommand{\M}{\mathcal{M}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\Gv}{\;\;\big|\;\;}
%\newcommand{\cP}{\mathcal{P}}
\newcommand{\proj}{\cP}
\newcommand{\pow}{\text{Pow}}
\newcommand{\sF}{\mathscr{F}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\sC}{\mathscr{C}}
\newcommand{\hJ}{\widehat{J}}
\newcommand{\bH}{\mathbf{H}}
\newcommand{\bM}{\mathbf{M}}
\newcommand{\tM}{\widetilde{M}}
\newcommand{\tE}{\widetilde{E}}
\newcommand{\hK}{\widehat{K}}
\newcommand{\hk}{\hat{k}}
\newcommand{\cN}{\mathcal{N}}
\newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}

@@

@@

\subsection{Sub-Path Sufficiency Principle}

@@

\WFcomment{This is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
\begin{align*}
  \sF_{k,\ell} &= \sF(M_{[k]},p_{[\ell]})\\
  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
  \sF_{1,1} \;\;\sub \cdots \sub\;\;
  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
  \underlabel_{\text{inference } d} \sF_{d,d}
\end{align*}

>>>>

\newcommand{\cM}{\mathcal{M}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\FDR}{\textnormal{FDR}}
\newcommand{\FCR}{\textnormal{FCR}}
\newcommand{\crt}{\phi}
\newcommand{\M}{\mathcal{M}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\Gv}{\;\;\big|\;\;}
%\newcommand{\cP}{\mathcal{P}}
\newcommand{\proj}{\cP}
\newcommand{\pow}{\text{Pow}}
\newcommand{\sF}{\mathscr{F}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\sC}{\mathscr{C}}
\newcommand{\hJ}{\widehat{J}}
\newcommand{\bH}{\mathbf{H}}
\newcommand{\bM}{\mathbf{M}}
\newcommand{\tM}{\widetilde{M}}
\newcommand{\tE}{\widetilde{E}}
\newcommand{\tV}{\widetilde{V}}
\newcommand{\tR}{\widetilde{R}}
\newcommand{\hK}{\widehat{K}}
\newcommand{\hk}{\hat{k}}
\newcommand{\cN}{\mathcal{N}}
\newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}

@@

Having discussed methods for constructing single-step $p$-values, we turn now to the problem of constructing a {\em stopping rule}; that is, an estimator $\hk$ of the completion index $k_0(Y)\in \{0,\ldots,d,\infty\}$. $M_k$ is ``rejected'' if and only if $\hk>k$. Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected. The number of type I errors, then, is $(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$. Define $V=(\hk-k_0)_+$, the number of excess rejections.

Note that the type I error $V$ is defined in a ``model-centric,'' as opposed to a ``variable-centric,'' fashion: in regression, our model-wise $V$ counts the number of true null {\em models} incorrectly rejected, instead of counting the number of noise variables included in the final model. We can define the (model-wise) familywise error rate (FWER) and false discovery rate (FDR) respectively as $\P_F(V >0)$ and $\E_F[V/(\hk \vee 1)]$.

We could have instead defined the ``variable-centric'' type I error $\tV$ as the number of noise variables incorrectly included in the final model, with $\tR$ denoting the total number of variables included. Here, by ``noise variable'' we mean one that has a non-zero coefficient in the full model. Using selective inference to control the variable-wise FWER $\P_\mu(\tV>0)$ and FDR $\E[\tV / (\tR \vee 1)]$ is an interesting topic for further study.

@@

We will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values, which have been studied in the literature by \WFcomment{literature review here.} We will focus on three such stopping rules: simple stop, proposed by \WFcomment{whom?}, and the more refined strong stop and forward stop, proposed by \citet{gsell2013sequential}.

While strong stop and forward stop tend to give more powerful stopping rules, they require independence among the $p$-values. Sections~\ref{sec:modelSPSP}--\ref{sec:pvalSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent. We will see that, typically, selected-model $p$-values are sequentially independent while saturated-model $p$-values are not.

\subsubsection{Basic Stop:}

\WFcomment{Is there a better name for this? Can we find it in the literature?}

The most obvious sequential procedure is to reject at each step until the first time $p_k > \alpha$, which we can formalize as
\[
\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
  p_k > \alpha\right\} - 1
\]
We will call this procedure ``basic stop.''

If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
\[
\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha
\]
\WFcomment{Trivial, but no doubt someone else proved this first...}

Selectively valid single-step $p$-values do not necessarily guarantee FWER control when $k_0$ is random. In that case, we need a bit more, but it is sufficient that we have type I error control for $p_k$ conditional on $k_0=k-1$; i.e., not just conditional on $M_{k-1}$ being a correct model, but given that $M_{k-1}$ is the {\em first} correct model.

\subsubsection{Strong Stop:}

\citet{gsell2013sequential} propose another method for controlling the model-wise FWER, which they called ``strong stop.'' Define
\[
  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
    \leq \frac{\alpha k}{d}\right\}
\]
Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, strong stop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.

\citet{gsell2013sequential} show that if $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones. Specifically, they require that
\[
\P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
\mid p_1, \ldots, p_{k_0}) \leq \prod_{\ell=k_0+1}^d \alpha_\ell
\]

\subsubsection{Forward Stop:}

\subsection{Independent $p$-Values and Sufficiency Principles}

\subsubsection{A Sub-Path Sufficiency Principle for SelectionAlgorithms}\label{sec:modelSPSP}

\subsubsection{A Sub-Path Sufficiency Principle for SelectionAlgorithms}\label{sec:modelSPSP}

@@

\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
\begin{align*}
  \sF_{k,\ell} &= \sF(M_{[k]},p_{[\ell]})\\
  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
  \sF_{1,1} \;\;\sub \cdots \sub\;\;
  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
  \underlabel_{\text{inference } d} \sF_{d,d}
\end{align*}
